---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r results = "hide"}
library(tidyverse)
library(knitr)
Sys.setlocale("LC_TIME","English")

dt <- read.csv("activity.csv", sep = ",")
```


## What is mean total number of steps taken per day?
1. Shows a histogram of the total number of steps taken each day
```{r}
dt_pday <- dt %>% group_by(date) %>% summarise(tot_steps = sum(steps, na.rm = TRUE))

ggplot(dt_pday, aes(tot_steps)) + geom_histogram(binwidth = 1000) +
  labs(title = "The total number of steps taken each day", x = "The total number of steps", y = "Frequency") +
  theme_bw()

mean_pday <- round(mean(dt_pday$tot_steps, na.rm = TRUE), 0)
median_pday <- median(dt_pday$tot_steps, na.rm = TRUE)
```
2. The mean of the total number of steps taken per day is `r mean_pday` and the median of the total number of steps taken per day is `r median_pday`.


## What is the average daily activity pattern?
1. Shows a time series plot of the 5-minute interval
```{r}
dt_pitv <- dt %>% group_by(interval) %>% summarise(avg_steps = mean(steps, na.rm = TRUE))

ggplot(dt_pitv, aes(x = interval, y = avg_steps)) + geom_line() +
  labs(title = "The average number of steps taken each interval", x = "Interval", y = "The average number of steps") +
  theme_bw()

max_pitv <- round(max(dt_pitv$avg_steps))
```
2. Amongst the 5-minute interval, `r max_pitv` contains the maximum number of steps.


## Imputing missing values
```{r}
na_steps <- sum(is.na(dt$steps))
```
1. The total number of missing values in the dataset is `r na_steps`.

2. Shows a histogram of the total number of steps taken each day.
```{r}
dt1 <- dt
dt1[is.na(dt1$steps), 1] <- 0

dt1_pday <- dt1 %>% group_by(date) %>% summarise(tot_steps = sum(steps, na.rm = TRUE))

ggplot(dt1_pday, aes(tot_steps)) + geom_histogram(binwidth = 1000) +
  labs(title = "The total number of steps taken each day", x = "The total number of steps", y = "Frequency") +
  theme_bw()

mean1_pday <- round(mean(dt1_pday$tot_steps, na.rm = TRUE), 0)
median1_pday <- median(dt1_pday$tot_steps, na.rm = TRUE)
```
3. The mean of the total number of steps taken per day is `r mean1_pday` and the median of the total number of steps taken per day is `r median1_pday`.


## Are there differences in activity patterns between weekdays and weekends?
1. Shows a panel plot containing a time series plot of the 5-minute interval by weekday
```{r}
dt1$date <- as.Date(as.character(dt1$date), format = "%Y-%m-%d")
dt1 <- dt1 %>% mutate(day = weekdays(date, abbreviate = TRUE),
                                weekday = case_when(day %in% c("Mon", "Tue", "Wed", "Thu", "Fri") ~ "weekday",
                                                    day %in% c("Sat", "Sun") ~ "weekend"))

dt1_pitv <- dt1 %>% group_by(weekday, interval) %>% summarise(avg_steps = mean(steps, na.rm = TRUE))

ggplot(dt1_pitv, aes(x = interval, y = avg_steps)) + geom_line() +
  labs(title = "The average number of steps taken each interval by day", x = "Interval", y = "The average number of steps") +
  facet_wrap(~ weekday) +
  theme_bw()
```

## Makes PA1_template.md and PA1_template.html files with knit2html() function
```{r}
knitr::knit2html(input = "PA1_template.Rmd", output = "PA1_template.md", force_v1 = TRUE)
```